### 2024-06-19 paper/paper.tex
mac

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks=true,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Splitting \texttt --------------------------------------------------

\let\textttOrig=\texttt
\def\texttt#1{\expandafter\textttOrig{\seqsplit{#1}}}
\renewcommand{\seqinsert}{\ifmmode
  \allowbreak
  \else\penalty6000\hspace{0pt plus 0.02em}\fi}

% --- Pandoc does not distinguish between links like [foo](bar) and
% --- [foo](foo) -- a simplistic Markdown model.  However, this is
% --- wrong:  in links like [foo](foo) the text is the url, and must
% --- be split correspondingly.
% --- Here we detect links \href{foo}{foo}, and also links starting
% --- with https://doi.org, and use path-like splitting (but not
% --- escaping!) with these links.
% --- Another vile thing pandoc does is the different escaping of
% --- foo and bar.  This may confound our detection.
% --- This problem we do not try to solve at present, with the exception
% --- of doi-like urls, which we detect correctly.

\makeatletter
\let\href@Orig=\href
\def\href@Urllike#1#2{\href@Orig{#1}{\begingroup
    \def\Url@String{#2}\Url@FormatString
    \endgroup}}
\def\href@Notdoi#1#2{\def\tempa{#1}\def\tempb{#2}%
  \ifx\tempa\tempb\relax\href@Urllike{#1}{#2}\else
  \href@Orig{#1}{#2}\fi}
\def\href#1#2{%
  \IfBeginWith{#1}{https://doi.org}%
  {\href@Urllike{#1}{#2}}{\href@Notdoi{#1}{#2}}}
\makeatother

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily D. Keitel, R. Tenorio, G. Ashton, R. Prix, (2021). \(ler\)
: LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers)
Rate calculator and
simulator. \textit{Journal of Open Source Software}, 6(60), 3000. \url{https://doi.org/10.21105/joss.03000}}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\texttt=\textttOrig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={ler : LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers) Rate calculator and simulator},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% --- We redefined \texttt, but in sections and captions we want the
% --- old definition
\let\addcontentslineOrig=\addcontentsline
\def\addcontentsline#1#2#3{\bgroup
  \let\texttt=\textttOrig\addcontentslineOrig{#1}{#2}{#3}\egroup}
\let\markbothOrig\markboth
\def\markboth#1#2{\bgroup
  \let\texttt=\textttOrig\markbothOrig{#1}{#2}\egroup}
\let\markrightOrig\markright
\def\markright#1{\bgroup
  \let\texttt=\textttOrig\markrightOrig{#1}\egroup}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{\(ler\) : LVK (LIGO-Virgo-KAGRA collaboration) Event
(compact-binary mergers) Rate calculator and simulator}

        \author[1]{Phurailatpam Hemantakumar}
          \author[2,3]{Anupreeta More}
          \author[4,5]{Harsh Narola}
          \author[1]{Ng Chung Yin (Leo)}
          \author[4,5]{Justin Janquart}
          \author[4,5]{Chris Van Den Broeck}
          \author[1]{Otto Akseli HANNUKSELA}
          \author[6]{Neha Singh}
          \author[6]{David Keitel}

      \affil[1]{Department of Physics, The Chinese University of Hong
Kong, Shatin, New Territories, Hong Kong}
      \affil[2]{The Inter-University Centre for Astronomy and
Astrophysics (IUCAA), Post Bag 4, Ganeshkhind, Pune 411007, India}
      \affil[3]{Kavli Institute for the Physics and Mathematics of the
Universe (IPMU), 5-1-5 Kashiwanoha, Kashiwa-shi, Chiba 277-8583, Japan}
      \affil[4]{Department of Physics, Utrecht University,
Princetonplein 1, 3584 CC Utrecht, The Netherlands}
      \affil[5]{Nikhef -- National Institute for Subatomic Physics,
Science Park, 1098 XG Amsterdam, The Netherlands}
      \affil[6]{Departament de F\textquotesingle isica, Universitat de
les Illes Balears, IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma,
Spain}
  \date{\vspace{-7ex}}

\begin{document}
\maketitle

\marginpar{

  \begin{flushleft}
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/10.21105/joss.03000}{\color{linky}{10.21105/joss.03000}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{https://github.com/openjournals/joss-reviews/issues/3000}{\color{linky}{Review}} \ExternalLink
    \item \href{https://github.com/pyfstat/pyfstat}{\color{linky}{Repository}} \ExternalLink
    \item \href{https://doi.org/10.5281/zenodo.4660591}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  \par\noindent\hrulefill\par

  \vspace{2mm}

  {\bfseries Editor:} \href{https://danielskatz.org/}{Daniel S. Katz} \ExternalLink \\
  \vspace{1mm}
    {\bfseries Reviewers:}
  \begin{itemize}
  \setlength\itemsep{0em}
    \item \href{https://github.com/khanx169}{@khanx169}
    \item \href{https://github.com/RobertRosca}{@RobertRosca}
    \end{itemize}
    \vspace{2mm}

  {\bfseries Submitted:} 26 January 2021\\
  {\bfseries Published:} 06 April 2021

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).

  \end{flushleft}
}

\section{Summary}\label{summary}

Gravitational waves (GWs) are ripples in the fabric of space and time
caused by the acceleration of unevenly distributed mass or masses.
Observable GWs are created especially during the violent cosmic events
of merging compact binaries, such as `binary black holes' (BBH), `binary
neutron stars' (BNS), and `neutron star and black hole pair' (NSBH). GWs
emitted by these events can be distorted or magnified by the
gravitational fields of massive objects such as galaxies or galaxy
clusters, a phenomenon known as gravitational lensing. Profound
comprehension of gravitational lensing's impact on GW signals is
imperative to their accurate interpretation and the extraction of
astrophysical insights therein. For this purpose, statistical modelling
of GWs lensing can provide valuable insights into the properties of the
lensing objects and GW sources. Such statistics require accurate and
efficient means to calculate the detectable lensing rates, which depend
on up-to-date modeling and implementation of lens and source properties
and their distribution. The outcomes of these computational analyses not
only contribute to generating dependable forecasts but also play an
important role in validating forthcoming lensing events (Janquart et al.
2023).

Obtaining precise outcomes in statistical analyses of this nature
necessitates the utilization of large-scale sampling, often numbering in
the millions. However, this process is computationally demanding. The
\(ler\) framework addresses this by employing innovative techniques to
optimize the workflow and computation efficiency required for handling
large-scale statistical analyses, essential for modeling detectable
events and calculating rates. Its integration of modular statistical
components enhances the framework's adaptability and extendability, thus
proving to be an invaluable asset in the evolving field of gravitational
wave research. Detailed description, source code, and examples are
available in \(ler\)
\href{https://ler.readthedocs.io/en/latest/}{documentation}.

\section{Statement of need}\label{statement-of-need}

\(ler\) is a statistical-based Python package specifically designed for
computing detectable rates of both lensed and unlensed gravitational
wave (GW) events, catering to the requirements of the LIGO-Virgo-KAGRA
Scientific Collaboration and astrophysics research scholars. The core
functionality of \(ler\) intricately hinges upon the interplay of
various components which include sampling the properties of
compact-binary sources, lens galaxies characteristics, solving lens
equations to derive properties of resultant images, and computing
detectable GW rates. This comprehensive functionality builds on the
leveraging of array operations and linear algebra from the \emph{numpy}
library, enhanced by interpolation methods from \emph{scipy} and
Python's \emph{multiprocessing} capabilities. Efficiency is further
boosted by the \emph{numba} library's Just-In-Time (\emph{njit})
compilation, optimizing extensive numerical computations and employing
the inverse transform sampling method to replace more cumbersome
rejection sampling. The modular design of \(ler\) not only optimizes
speed and functionality but also ensures adaptability and upgradability,
supporting the integration of additional statistics as research evolves.
Currently, \(ler\) is an important tool in generating simulated
gravitational wave (GW) events---both lensed and unlensed---and provides
astrophysically accurate distributions of event-related parameters for
both detectable and non-detectable events. This functionality aids in
event validation and enhances the forecasting of detection capabilities
across various GW detectors to study such events. The architecture of
the \(ler\) API facilitates seamless compatibility with other software
packages, allowing researchers to integrate and utilize its
functionalities based on specific scientific requirements.

\section{Design and Structure}\label{design-and-structure}

The architecture of the \(ler\) API is deliberately organized such that
each distinct functionality holds its own significance in scientific
research. Simultaneously, these functionalities seamlessly integrate and
can be employed collectively to accommodate diverse scientific
objectives. Key features of \(ler\) and its dependencies can be
summarized as follows:

\begin{itemize}
\tightlist
\item
  Sampling Gravitational Wave (GW) Source Properties:

  \begin{itemize}
  \tightlist
  \item
    For the unlensed events, The sampling distribution
    \((\,R_m(z_s)\,)\) for the source's redshift \((\,z_s\,)\) is
    derived from the merger rate density of compact binaries, which, in
    turn, is based on the star formation rate. The code is meticulously
    designed to enable the straightforward integration of future updates
    or user-specified distributions of these sources.
  \item
    The process of sampling both intrinsic and extrinsic parameters of
    GW sources, represented by \(\theta\), utilizes the prior
    distributions \((P(\theta))\) available within the \emph{gwcosmo}
    and \emph{bilby} Python packages. Additionally, users have the
    option to initialize the package with custom distributions of their
    choosing.
  \end{itemize}
\item
  Sampling of lens galaxies attributes and source red-shifts:

  \begin{itemize}
  \tightlist
  \item
    For lensed case, the source redshift \((z_s)\) is sampled under the
    strong lensing condition \((\text{SL})\), based on the precomputed
    probability of strong lensing with source at \(z_s\)
    \((\text{optical depth: }P\left(\text{SL}|z_s\right)\,)\). This
    probability can be recalculated for specified configurations of lens
    galaxies, leveraging \emph{multiprocessing} and \emph{njit}
    functionalities for enhanced efficiency.
  \item
    The package utilizes the Elliptical Power Law with external shear
    (EPL+Shear) model (Wempe et al. 2022) for galaxy parameters
    \((\theta_L)\) sampling, following reference
    (https://arxiv.org/abs/2106.06303)(Wierda et al. 2021). Rejection
    sampling is applied on the above samples on condition that whether
    the event is strongly lensed or not,
    \(P\left(\text{SL}|z_s,\theta_L\right)\).
  \end{itemize}
\item
  Generation of image properties:

  \begin{itemize}
  \tightlist
  \item
    Source position \((\beta)\) is sampled from the caustic in the
    source plane.
  \item
    Sampled lens properties and source position is fed in
    \emph{lenstronomy} to generate properties of the images. This is the
    slowest part of the entire simulation, which \(ler\) tackles through
    parallelization with multiprocessing.
  \item
    Image properties like magnification \((\mu_i)\) and time delay
    (\(\Delta t_i\)) modifies the original source signal strength,
    changing the signal-to-noise ratio SNR and our ability to detect.
  \end{itemize}
\item
  Calculation of Detectable Merger Rates Per Year:

  \begin{itemize}
  \tightlist
  \item
    The calculation of rates necessitates integration over simulated
    events that meet specific detection criteria. This process includes
    computing SNRs \((\rho)\) for each event or its lensed images,
    followed by an assessment against a predetermined threshold(s)
    \((\rho_{th})\).
  \item
    SNR calculations are optimized using
    \href{https://github.com/hemantaph/gwsnr}{\emph{gwsnr}}, leveraging
    interpolation and multiprocessing for accuracy and speed.
  \item
    Simulated events and rate results, along with input configurations,
    are systematically archived for easy access and future analysis.
    Additionally, all interpolators used in the process are preserved
    for future applications.
  \end{itemize}
\end{itemize}

\section{Equations}\label{equations}

\(\textbf{Detectable Unlensed rates:}\)

\begin{equation*}
\begin{split}
R_U = \int & dz_s \frac{dV_c}{dz_s}\frac{R_m(z_s)}{1+z_s}\left\{\Theta[\rho(z_s,\theta)-\rho_{th}] P(\theta) d\theta \right\}
\end{split}
\end{equation*}

\(z_s\): GW source redshift, \(\frac{dV_c}{dz_s}\): Differential
co-moving volume, \(\frac{1}{1+z_s}\): Time dilation correction factor,
\(R_m(z_s)\): source frame merger rate density, \(\theta\): GW source
parameters, \(P\): probability distribution, \(\rho\): SNR,
\(\rho_{th}\): SNR threshold, \(\Theta\): Heaviside function to select
detectable GW events.

\(\textbf{Detectable Lensed rates:}\)

\begin{equation*}
\begin{split}
R_L = \int & dz_s \frac{dV_c}{dz_s}\tau(z_s)\frac{R_m(z_s)}{1+z_s} \,\mathcal{O}_{images}(z_s,\theta,\mu_i,\Delta t_i, \rho_{th}) \, \\ 
& \, P(\theta) P(\theta_L|\text{SL},z_s) P(\beta|\text{SL}) d\theta d\beta d\theta_L dz_s 
\end{split}
\end{equation*}

\(\tau(z_s)\): Optical-depth of strong lensing, \(\theta_L\): lens
parameters, \(\beta\): source position, \(\mu\): image magnification,
\(\Delta t\): image time delay, \(\mathcal{O}\): operator to select
dectectable lensed GW events, \(i\): index of images of a lensed event,
\(\text{SL}\): strong lensing condition.

\section{Acknowledgements}\label{acknowledgements}

The authors express their sincere appreciation for the significant
contributions that have been instrumental in completing this research.
Special thanks are extended to the academic advisors for their
invaluable guidance and steadfast support. The collaborative efforts and
enriching discussions with research colleagues significantly enhanced
the study's quality. Acknowledgement is given to the Department of
Physics, The Chinese University of Hong Kong, for the Postgraduate
Studentship that facilitated this research. Further gratitude is
extended to the Netherlands Organisation for Scientific Research (NWO)
for their support. The authors also recognize the contributions of
individuals who added empirical depth to this work. Appreciation is
conveyed for the computational resources provided by the LIGO
Laboratory, supported by National Science Foundation Grants
No.~PHY-0757058 and No.~PHY-0823459.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Janquart2023}
Janquart, J, M Wright, S Goyal, J C L Chan, A Ganguly, Á Garrón, D
Keitel, et al. 2023. {``Follow-up Analyses to the O3 LIGO--Virgo--KAGRA
Lensing Searches.''} \emph{Monthly Notices of the Royal Astronomical
Society} 526 (3): 3832--60.
\url{https://doi.org/10.1093/mnras/stad2909}.

\bibitem[\citeproctext]{ref-Wempe2022}
Wempe, Ewoud, Léon V. E. Koopmans, A. Renske A. C. Wierda, Otto Akseli
Hannuksela, Alberto Agnello, Cyril Bonvin, Bendetta Bucciarelli, et al.
2022. {``A Lensing Multi-Messenger Channel: Combining LIGO-Virgo-Kagra
Lensed Gravitational-Wave Measurements with Euclid Observations.''}
\url{https://arxiv.org/abs/2204.08732}.

\bibitem[\citeproctext]{ref-Wierda2021}
Wierda, A. Renske A. C., Ewoud Wempe, Otto A. Hannuksela, Léon V. E.
Koopmans, Alberto Agnello, Cyril Bonvin, Bendetta Bucciarelli, et al.
2021. {``Beyond the Detector Horizon: Forecasting Gravitational-Wave
Strong Lensing.''} \emph{The Astrophysical Journal} 921 (1): 154.
\url{https://doi.org/10.3847/1538-4357/ac1bb4}.

\end{CSLReferences}

\end{document}

### 2024-06-20 paper/paper.tex
mac

@@
 leveraging of array operations and linear algebra from the \emph{numpy}
-library, enhanced by interpolation methods from \emph{scipy} and
-Python's \emph{multiprocessing} capabilities. Efficiency is further
-boosted by the \emph{numba} library's Just-In-Time (\emph{njit})
+(NumPy Community 2022) library, enhanced by interpolation methods from
+\emph{scipy} (Virtanen et al. 2020) and Python's \emph{multiprocessing}
+capabilities. Efficiency is further boosted by the \emph{numba} (Lam,
+Pitrou, and Seibert 2022) library's Just-In-Time (\emph{njit})
 compilation, optimizing extensive numerical computations and employing
@@
   \item
-    The process of sampling both intrinsic and extrinsic parameters of
-    GW sources, represented by \(\theta\), utilizes the prior
-    distributions \((P(\theta))\) available within the \emph{gwcosmo}
-    and \emph{bilby} Python packages. Additionally, users have the
-    option to initialize the package with custom distributions of their
-    choosing.
+    Intrinsic and extrinsic parameters \((\theta)\) of GW sources are
+    sampled using prior distributions \((P(\theta))\) from the gwcosmo
+    (gwcosmo Contributors 2022) and bilby (\textbf{Ashton2019?})
+    packages, with options for users to input custom distributions.
   \end{itemize}
@@
     probability of strong lensing with source at \(z_s\)
-    \((\text{optical depth: }P\left(\text{SL}|z_s\right)\,)\). This
-    probability can be recalculated for specified configurations of lens
-    galaxies, leveraging \emph{multiprocessing} and \emph{njit}
+    \((\text{optical depth: }P\left(\text{SL}|z_s\right) \text{ or }\tau(z_s)\,)\).
+    This probability can be recalculated for specified configurations of
+    lens galaxies, leveraging \emph{multiprocessing} and \emph{njit}
     functionalities for enhanced efficiency.
@@
     (EPL+Shear) model (Wempe et al. 2022) for galaxy parameters
-    \((\theta_L)\) sampling, following reference
-    (https://arxiv.org/abs/2106.06303)(Wierda et al. 2021). Rejection
+    \((\theta_L)\) sampling, following Wierda et al. (2021). Rejection
     sampling is applied on the above samples on condition that whether
@@
     Sampled lens properties and source position is fed in
-    \emph{lenstronomy} to generate properties of the images. This is the
-    slowest part of the entire simulation, which \(ler\) tackles through
-    parallelization with multiprocessing.
+    \emph{lenstronomy} (Birrer et al. 2021) to generate properties of
+    the images. This is the slowest part of the entire simulation, which
+    \(ler\) tackles through parallelization with multiprocessing.
   \item
@@
     for future applications.
+  \item
+    Most cosmology-related calculations within the \emph{ler} package
+    are performed using the \emph{astropy} library (Astropy
+    Collaboration 2013). The default cosmological model is LambdaCDM
+    (\(H_0=70,\, \Omega_m=0.3,\, \Omega_\Lambda=0.7\)); however, users
+    have the flexibility to employ any cosmology available in
+    \emph{astropy}. All internal calculations in \emph{ler} will then be
+    based on the user-selected cosmological model.
   \end{itemize}
@@
 
+\[
 \begin{equation*}
@@
 \end{equation*}
+\]
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-astropy}
+Astropy Collaboration. 2013. {``{Astropy: A community Python package for
+astronomy}.''} \emph{Astronomy and Astrophysics} 558 (October).
+\url{https://doi.org/10.1051/0004-6361/201322068}.
+
+\bibitem[\citeproctext]{ref-Birrer2021}
+Birrer, Simon, Anowar J. Shajib, Daniel Gilman, Aymeric Galan, Jelle
+Aalbers, Martin Millon, Robert Morgan, et al. 2021. {``Lenstronomy II: A
+Gravitational Lensing Software Ecosystem.''} \emph{Journal of Open
+Source Software} 6 (62): 3283.
+\url{https://doi.org/10.21105/joss.03283}.
+
+\bibitem[\citeproctext]{ref-gwcosmo}
+gwcosmo Contributors. 2022. {``{gwcosmo: A Python package for
+gravitational-wave cosmology}.''} \emph{GitHub Repository}. GitHub.
+\url{https://github.com/gwcosmo/gwcosmo}.
+
 \bibitem[\citeproctext]{ref-Janquart2023}
 Janquart, J, M Wright, S Goyal, J C L Chan, A Ganguly, Á Garrón, D
-Keitel, et al. 2023. {``Follow-up Analyses to the O3 LIGO--Virgo--KAGRA
-Lensing Searches.''} \emph{Monthly Notices of the Royal Astronomical
+Keitel, et al. 2023. {``{Follow-up analyses to the O3 LIGO--Virgo--KAGRA
+lensing searches}.''} \emph{Monthly Notices of the Royal Astronomical
 Society} 526 (3): 3832--60.
 \url{https://doi.org/10.1093/mnras/stad2909}.
+
+\bibitem[\citeproctext]{ref-numba}
+Lam, Stan, Stéphane Pitrou, and Mark Seibert. 2022. {``Numba: A High
+Performance Python Compiler.''} \emph{Numba Documentation}. Anaconda,
+Inc. \url{https://numba.pydata.org/}.
+
+\bibitem[\citeproctext]{ref-numpy}
+NumPy Community. 2022. {``NumPy: A Fundamental Package for Scientific
+Computing with Python.''} \emph{NumPy Website}. NumPy.
+\url{https://numpy.org/}.
+
+\bibitem[\citeproctext]{ref-scipy}
+Virtanen, Pauli, Ralf Gommers, Travis E. Oliphant, Matt Haberland, Tyler
+Reddy, David Cournapeau, Evgeni Burovski, et al. 2020. {``{SciPy 1.0:
+Fundamental Algorithms for Scientific Computing in Python}.''}
+\emph{Nature Methods}. SciPy. \url{https://www.scipy.org/}.
 

### 2024-06-20 paper/paper.tex
mac

@@
     sampled using prior distributions \((P(\theta))\) from the gwcosmo
-    (gwcosmo Contributors 2022) and bilby (\textbf{Ashton2019?})
-    packages, with options for users to input custom distributions.
+    (gwcosmo Contributors 2022) and bilby (Ashton et al. 2019) packages,
+    with options for users to input custom distributions.
   \end{itemize}
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Ashton2019}
+Ashton, Gregory, Moritz Hübner, Paul D. Lasky, Colm Talbot, Kendall
+Ackley, Sylvia Biscoveanu, Qi Chu, et al. 2019. {``Bilby: A
+User-Friendly Bayesian Inference Library for Gravitational-Wave
+Astronomy.''} \emph{The Astrophysical Journal Supplement Series} 241
+(2): 27. \url{https://doi.org/10.3847/1538-4365/ab06fc}.
+
 \bibitem[\citeproctext]{ref-astropy}

### 2024-06-22 paper/paper.tex
mac

@@
 
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily D. Keitel, R. Tenorio, G. Ashton, R. Prix, (2021). \(ler\)
+\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Phurailatpam Hemantakumar, Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli HANNUKSELA, Neha Singh, and David Keitel, (2024).
+\(ler\)
 : LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers)
 Rate calculator and
-simulator. \textit{Journal of Open Source Software}, 6(60), 3000. \url{https://doi.org/10.21105/joss.03000}}}}
+simulator. \textit{Journal of Open Source Software}, (), . \url{https://doi.org/10.xxxxxx/draft}}}}
 
@@
 
-  {\bfseries DOI:} \href{https://doi.org/10.21105/joss.03000}{\color{linky}{10.21105/joss.03000}}
+  {\bfseries DOI:} \href{https://doi.org/10.xxxxxx/draft}{\color{linky}{10.xxxxxx/draft}}
 
@@
     \setlength\itemsep{0em}
-    \item \href{https://github.com/openjournals/joss-reviews/issues/3000}{\color{linky}{Review}} \ExternalLink
-    \item \href{https://github.com/pyfstat/pyfstat}{\color{linky}{Repository}} \ExternalLink
-    \item \href{https://doi.org/10.5281/zenodo.4660591}{\color{linky}{Archive}} \ExternalLink
+    \item \href{}{\color{linky}{Review}} \ExternalLink
+    \item \href{https://github.com/hemantaph/ler}{\color{linky}{Repository}} \ExternalLink
+    \item \href{https://zenodo.org/badge/latestdoi/626733473}{\color{linky}{Archive}} \ExternalLink
   \end{itemize}
@@
 
-  {\bfseries Editor:} \href{https://danielskatz.org/}{Daniel S. Katz} \ExternalLink \\
+  {\bfseries Editor:} \href{}{Pending Editor} \ExternalLink \\
   \vspace{1mm}
@@
   \setlength\itemsep{0em}
-    \item \href{https://github.com/khanx169}{@khanx169}
-    \item \href{https://github.com/RobertRosca}{@RobertRosca}
+    \item \href{https://github.com/@Pending Reviewer}{@@Pending Reviewer}
+    \item \href{https://github.com/}{@}
     \end{itemize}
@@
 
-  {\bfseries Submitted:} 26 January 2021\\
-  {\bfseries Published:} 06 April 2021
+  {\bfseries Submitted:} 19 June 2024\\
+  {\bfseries Published:} 19 June 2024
 
@@
     SNR calculations are optimized using
-    \href{https://github.com/hemantaph/gwsnr}{\emph{gwsnr}}, leveraging
-    interpolation and multiprocessing for accuracy and speed.
+    \href{https://gwsnr.readthedocs.io/en/latest/}{\emph{gwsnr}},
+    leveraging interpolation and multiprocessing for accuracy and speed.
   \item
@@
 
-\[
 \begin{equation*}
@@
 \end{equation*}
-\]
 

### 2024-06-22 ler/lens_galaxy_population/lens_galaxy_parameter_distribution.py

@@ -294,7 +294,7 @@ class LensGalaxyParameterDistribution(CBCSourceParameterDistribution, ImagePrope
         # To find the normalization constant of the pdf p(z)
         # this under the assumption that the event is strongly lensed
         # Define the merger-rate density function
-        pdf_unnormalized_ = lambda z: self.merger_rate_density_src_frame(np.array([z])) * self.strong_lensing_optical_depth(np.array([z]))
+        pdf_unnormalized_ = lambda z: self.merger_rate_density_src_frame(np.array([z]), param=self.merger_rate_density_param) * self.strong_lensing_optical_depth(np.array([z]))
         pdf_unnormalized = lambda z: pdf_unnormalized_(z)[0]
 
         self.normalization_pdf_z_lensed = quad(

### 2024-06-22 ler/utils/utils.py

@@ -187,16 +187,24 @@ def rejection_sample(pdf, xmin, xmax, size=100, chunk_size=10000):
     """
     x = np.linspace(xmin, xmax, chunk_size)
     y = pdf(x)
+    # Maximum value of the pdf
     ymax = np.max(y)
+
     # Rejection sample in chunks
     x_sample = []
     while len(x_sample) < size:
         x_try = np.random.uniform(xmin, xmax, size=chunk_size)
         pdf_x_try = pdf(x_try) # Calculate the pdf at the random x values
+        # this is for comparing with the pdf value at x_try, will be used to accept or reject the sample
         y_try = np.random.uniform(0, ymax, size=chunk_size)
-        ymax = max(ymax, np.max(pdf_x_try))  # Update the maximum value of the pdf
+        
+        # Update the maximum value of the pdf
+        ymax = max(ymax, np.max(pdf_x_try))  
+        
+        # applying condition to accept the sample
         # Add while retaining 1D shape of the list
         x_sample += list(x_try[y_try < pdf_x_try])
+
     # Transform the samples to a 1D numpy array
     x_sample = np.array(x_sample).flatten()
     # Return the correct number of samples
@@ -234,6 +242,7 @@ def rejection_sample2d(pdf, xmin, xmax, ymin, ymax, size=100, chunk_size=10000):
     x = np.random.uniform(xmin, xmax, chunk_size)
     y = np.random.uniform(ymin, ymax, chunk_size)
     z = pdf(x, y)
+    # Maximum value of the pdf
     zmax = np.max(z)
 
     # Rejection sample in chunks
@@ -242,12 +251,15 @@ def rejection_sample2d(pdf, xmin, xmax, ymin, ymax, size=100, chunk_size=10000):
     while len(x_sample) < size:
         x_try = np.random.uniform(xmin, xmax, size=chunk_size)
         y_try = np.random.uniform(ymin, ymax, size=chunk_size)
-
+        pdf_xy_try = pdf(x_try, y_try)
+        # this is for comparing with the pdf value at x_try, will be used to accept or reject the sample
         z_try = np.random.uniform(0, zmax, size=chunk_size)
-        zmax = max(zmax, np.max(z_try))
+        
+        # Update the maximum value of the pdf
+        zmax = max(zmax, np.max(pdf_xy_try))
 
-        x_sample += list(x_try[z_try < pdf(x_try, y_try)])
[…]

### 2024-06-22 ler/gw_source_population/cbc_source_redshift_distribution.py
mac

@@ -221,16 +221,18 @@ class CBCSourceRedshiftDistribution(object):
             if self.event_type == "BBH":
                 R0 = 23.9 * 1e-9
             elif self.event_type == "BNS":
-                R0 = 150.5 * 1e-9
+                R0 = 105.5 * 1e-9
             elif self.event_type == "NSBH":
                 R0 = 45.0 * 1e-9
             else:
                 raise ValueError("event_type must be one of 'BBH', 'BNS', 'NSBH'")
-            self.merger_rate_density_param = dict(R0=R0, b2=1.6, b3=2.0, b4=30)
+            merger_rate_density_param = dict(R0=R0, b2=1.6, b3=2.0, b4=30)
+        self.merger_rate_density_param = merger_rate_density_param
+
 
         # To find the normalization constant of the pdf p(z)
         # merger_rate_density_src_frame is njit function and takes array as input but quad integrand function takes only float as input
-        merger_rate_density_src_frame = lambda z: self.merger_rate_density_src_frame(np.array([z]))[0]
+        merger_rate_density_src_frame = lambda z: self.merger_rate_density_src_frame(zs=np.array([z]), param=merger_rate_density_param)[0]
         # this normalization is important to find the correct pdf p(z)
         self.normalization_pdf_z = quad(
             merger_rate_density_src_frame,
@@ -380,6 +382,7 @@ class CBCSourceRedshiftDistribution(object):
             b3 = param["b3"]
             b4 = param["b4"]
 
+        #print(f"\nR0: {R0}\n")
         return merger_rate_density_bbh_popI_II_oguri2018(zs=zs, R0=R0, b2=b2, b3=b3, b4=b4)
     
     def star_formation_rate_madau_dickinson2014(self,
@@ -647,7 +650,7 @@ class CBCSourceRedshiftDistribution(object):
     
     @merger_rate_density.setter
     def merger_rate_density(self, merger_rate_density):
-        error_msg = ValueError(f"merger_rate_density must be one of {self.merger_rate_density_model_list}")
+        # error_msg = ValueError(f"merger_rate_density must be one of {self.merger_rate_density_model_list}")
         # check if it is a string
         if isinstance(merger_rate_density, str):
             try:

### 2024-06-25 ler/gw_source_population/cbc_source_redshift_distribution.py
mac

@@ -86,7 +86,7 @@ class CBCSourceRedshiftDistribution(object):
     |                                     | and its parameters               |
     +-------------------------------------+----------------------------------+
     |:attr:`~sample_source_redshift`      | Function to sample source        |
-    |                                     | redshifts (source frame)         |
+    |                                     | redshifts                        |
     +-------------------------------------+----------------------------------+
 
     Instance Methods
@@ -104,7 +104,7 @@ class CBCSourceRedshiftDistribution(object):
     |:meth:`~pdf_z`                       | Function to compute the pdf      |
     |                                     | p(z)                             |
     +-------------------------------------+----------------------------------+
-    |:meth:`~merger_rate_density_src_frame`                                  |
+    |:meth:`~merger_rate_density_detector_frame`                                  |
     +-------------------------------------+----------------------------------+
     |                                     | Function to compute the merger   |
     |                                     | rate density (source frame)      |
@@ -231,11 +231,11 @@ class CBCSourceRedshiftDistribution(object):
 
 
         # To find the normalization constant of the pdf p(z)
-        # merger_rate_density_src_frame is njit function and takes array as input but quad integrand function takes only float as input
-        merger_rate_density_src_frame = lambda z: self.merger_rate_density_src_frame(zs=np.array([z]), param=merger_rate_density_param)[0]
+        # merger_rate_density_detector_frame is njit function and takes array as input but quad integrand function takes only float as input
+        merger_rate_density_detector_frame = lambda z: self.merger_rate_density_detector_frame(zs=np.array([z]), param=merger_rate_density_param)[0]
         # this normalization is important to find the correct pdf p(z)
         self.normalization_pdf_z = quad(
-            merger_rate_density_src_frame,
+            merger_rate_density_detector_frame,
             z_min,
             z_max
         )[0]
@@ -299,9 +299,9 @@ class CBCSourceRedshiftDistribution(object):
         >>> pdf = cbc.pdf_z(zs=0.1)
         """
 
-        return self.merger_rate_density_src_frame(zs=zs,param=param)/self.normalization_pdf_z
+        return self.merger_rate_density_detector_frame(zs=zs,param=param)/self.normalization_pdf_z
 
-    def merger_rate_density_src_frame(self, zs, param=None):
+    def merger_rate_density_detector_frame(self, zs, param=None):
         """
[…]

### 2024-06-25 ler/gw_source_population/jit_functions.py
mac

@@ -143,7 +143,7 @@ def star_formation_rate_madau_dickinson2014(
     return 0.015 * (1 + zs) ** af / (1 + ((1 + zs) / cf) ** bf)
 
 
-@jit
+# @jit
 def merger_rate_density_bbh_primordial_ken2022(
         zs, cosmology=cosmo, n0=0.044 * 1e-9, t0=13.786885302009708
     ):

### 2024-06-26 paper/paper.tex
mac

@@
 
-\(ler\) is a statistical-based Python package specifically designed for
-computing detectable rates of both lensed and unlensed gravitational
-wave (GW) events, catering to the requirements of the LIGO-Virgo-KAGRA
-Scientific Collaboration and astrophysics research scholars. The core
-functionality of \(ler\) intricately hinges upon the interplay of
-various components which include sampling the properties of
+\(ler\) is a statistics-based Python package specifically designed for
+computing detectable rates of both lensed and unlensed GW events,
+catering to the requirements of the LIGO-Virgo-KAGRA Scientific
+Collaboration (The LIGO Scientific Collaboration et al. 2015) (Acernese
+et al. 2014) (Akutsu et al. 2020) and astrophysics research scholars.
+The core functionality of \(ler\) intricately hinges upon the interplay
+of various components which include sampling the properties of
 compact-binary sources, lens galaxies characteristics, solving lens
@@
 supporting the integration of additional statistics as research evolves.
-Currently, \(ler\) is an important tool in generating simulated
-gravitational wave (GW) events---both lensed and unlensed---and provides
-astrophysically accurate distributions of event-related parameters for
-both detectable and non-detectable events. This functionality aids in
-event validation and enhances the forecasting of detection capabilities
-across various GW detectors to study such events. The architecture of
-the \(ler\) API facilitates seamless compatibility with other software
-packages, allowing researchers to integrate and utilize its
-functionalities based on specific scientific requirements.
+Currently, \(ler\) is an important tool in generating simulated GW
+events---both lensed and unlensed---and provides astrophysically
+accurate distributions of event-related parameters for both detectable
+and non-detectable events. This functionality aids in event validation
+and enhances the forecasting of detection capabilities across various GW
+detectors to study such events. The architecture of the \(ler\) API
+facilitates seamless compatibility with other software packages,
+allowing researchers to integrate and utilize its functionalities based
+on specific scientific requirements.
 
@@
 \item
-  Sampling Gravitational Wave (GW) Source Properties:
+  Sampling GW Source Properties:
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-VIRGO2015}
+Acernese, F, M Agathos, K Agatsuma, D Aisa, N Allemandou, A Allocca, J
+Amarni, et al. 2014. {``Advanced Virgo: A Second-Generation
+Interferometric Gravitational Wave Detector.''} \emph{Classical and
+Quantum Gravity} 32 (2): 024001.
+\url{https://doi.org/10.1088/0264-9381/32/2/024001}.
+
+\bibitem[\citeproctext]{ref-KAGRA2021}
+Akutsu, T, M Ando, K Arai, Y Arai, S Araki, A Araya, N Aritomi, et al.
+2020. {``{Overview of KAGRA: Detector design and construction
+history}.''} \emph{Progress of Theoretical and Experimental Physics}
+2021 (5): 05A101. \url{https://doi.org/10.1093/ptep/ptaa125}.
+
 \bibitem[\citeproctext]{ref-Ashton2019}
@@
 
+\bibitem[\citeproctext]{ref-LIGO2015}
+The LIGO Scientific Collaboration, J Aasi, B P Abbott, R Abbott, T
+Abbott, M R Abernathy, K Ackley, et al. 2015. {``Advanced LIGO.''}
+\emph{Classical and Quantum Gravity} 32 (7): 074001.
+\url{https://doi.org/10.1088/0264-9381/32/7/074001}.
+
 \bibitem[\citeproctext]{ref-scipy}

### 2024-06-27 ler/lens_galaxy_population/lens_galaxy_parameter_distribution.py
mac

@@ -294,7 +294,7 @@ class LensGalaxyParameterDistribution(CBCSourceParameterDistribution, ImagePrope
         # To find the normalization constant of the pdf p(z)
         # this under the assumption that the event is strongly lensed
         # Define the merger-rate density function
-        pdf_unnormalized_ = lambda z: self.merger_rate_density_src_frame(np.array([z]), param=self.merger_rate_density_param) * self.strong_lensing_optical_depth(np.array([z]))
+        pdf_unnormalized_ = lambda z: self.merger_rate_density_detector_frame(np.array([z]), param=self.merger_rate_density_param) * self.strong_lensing_optical_depth(np.array([z]))
         pdf_unnormalized = lambda z: pdf_unnormalized_(z)[0]
 
         self.normalization_pdf_z_lensed = quad(

### 2024-06-27 ler/rates/ler.py

@@ -300,7 +300,7 @@ class LeR(LensGalaxyParameterDistribution):
         print("npool = ", self.npool)
         print("z_min = ", self.z_min)
         print("z_max = ", self.z_max)
-        print("event_type = ", self.event_type)
+        print(f"event_type = '{self.event_type}'")
         print("size = ", self.size)
         print("batch_size = ", self.batch_size)
         print("cosmology = ", self.cosmo)
@@ -320,7 +320,7 @@ class LeR(LensGalaxyParameterDistribution):
         print("create_new_interpolator=", self.gw_param_sampler_dict["create_new_interpolator"])
 
         print("\n LeR also takes LensGalaxyParameterDistribution params as kwargs, as follows:")
-        print("lens_type = ", self.gw_param_sampler_dict["lens_type"])
+        print(f"lens_type = '{self.gw_param_sampler_dict['lens_type']}'")
         print("lens_functions = ", self.gw_param_sampler_dict["lens_functions"])
         print("lens_priors = ", self.gw_param_sampler_dict["lens_priors"])
         print("lens_priors_params = ", self.gw_param_sampler_dict["lens_priors_params"])
@@ -341,9 +341,9 @@ class LeR(LensGalaxyParameterDistribution):
             print("mtot_resolution = ", self.snr_calculator_dict["mtot_resolution"])
             print("ratio_resolution = ", self.snr_calculator_dict["ratio_resolution"])
             print("sampling_frequency = ", self.snr_calculator_dict["sampling_frequency"])
-            print("waveform_approximant = ", self.snr_calculator_dict["waveform_approximant"])
+            print(f"waveform_approximant = '{self.snr_calculator_dict['waveform_approximant']}'")
             print("minimum_frequency = ", self.snr_calculator_dict["minimum_frequency"])
-            print("snr_type = ", self.snr_calculator_dict["snr_type"])
+            print(f"snr_type = '{self.snr_calculator_dict['snr_type']}'")
             print("psds = ", self.snr_calculator_dict["psds"])
             print("ifos = ", self.snr_calculator_dict["ifos"])
             print("interpolator_dir = ", self.snr_calculator_dict["interpolator_dir"])
@@ -354,54 +354,54 @@ class LeR(LensGalaxyParameterDistribution):
         del self.gwsnr
 
         print("\n For reference, the chosen source parameters are listed below:")
[…]

### 2024-06-27 ler/gw_source_population/cbc_source_redshift_distribution.py
mac

@@ -107,7 +107,7 @@ class CBCSourceRedshiftDistribution(object):
     |:meth:`~merger_rate_density_detector_frame`                                  |
     +-------------------------------------+----------------------------------+
     |                                     | Function to compute the merger   |
-    |                                     | rate density (source frame)      |
+    |                                     | rate density (detector frame)    |
     +-------------------------------------+----------------------------------+
     |:meth:`~create_lookup_table`         | Function to create a lookup      |
     |                                     | table for the differential       |
@@ -275,7 +275,7 @@ class CBCSourceRedshiftDistribution(object):
 
     def pdf_z(self, zs, param=None):
         """
-        Function to compute the pdf p(z). The output is in source frame and is normalized.
+        Function to compute the pdf p(z). The output is in detector frame and is normalized.
 
         Parameters
         ----------
@@ -303,7 +303,7 @@ class CBCSourceRedshiftDistribution(object):
 
     def merger_rate_density_detector_frame(self, zs, param=None):
         """
-        Function to compute the merger rate density (source frame). The output is in source frame and is unnormalized.
+        Function to compute the merger rate density (detector frame). The output is in detector frame and is unnormalized.
 
         Parameters
         ----------
@@ -318,7 +318,7 @@ class CBCSourceRedshiftDistribution(object):
         ----------
         rate_density : `numpy.ndarray`
             1D array of floats
-            merger rate density (source frame) (Mpc^-3 yr^-1)
+            merger rate density (detector frame) (Mpc^-3 yr^-1)
 
         Examples
         ----------
@@ -341,7 +341,7 @@ class CBCSourceRedshiftDistribution(object):
         zs, R0=23.9 * 1e-9, b2=1.6, b3=2.0, b4=30, param=None
     ):
         """
-        Function to compute the merger rate density (PopI/PopII). Reference: Oguri et al. (2018). The output is in detector frame and is unnormalized.
+        Function to compute the merger rate density (PopI/PopII). Reference: Oguri et al. (2018). The output is in source frame and is unnormalized.
 
         Parameters
         ----------
@@ -604,7 +604,7 @@ class CBCSourceRedshiftDistribution(object):
     @property
     def sample_source_redshift(self):
         """
-        Function to sample source redshifts (source frame) between z_min and z_max from the source galaxy population
+        Function to sample source redshifts (detector frame) between z_min and z_max from the detector galaxy population
 
         Parameters
         ----------

### 2024-06-30 ler/rates/ler.py
mac

@@ -1580,15 +1580,39 @@ class LeR(LensGalaxyParameterDistribution):
             print(f"total unlensed rate (yr^-1): {total_rate}")
 
         print(f"storing detectable unlensed params in {output_path}")
+        print(f"storing meta data in {meta_data_path}")
 
         if trim_to_size:
             # trim the final param dictionary
             print(f"\n trmming final result to size={size}")
             param_final = get_param_from_json(output_path)
+            # randomly select size number of samples
+            len_ = len(list(param_final.values())[0])
+            idx = np.random.choice(len_, size, replace=False)
             # trim the final param dictionary, randomly, without repeating
             for key, value in param_final.items():
-                param_final[key] = param_final[key][:size]
+                param_final[key] = value[idx]
 
+            # change meta data
+            meta_data = load_json(meta_data_path)
+            old_events_total = meta_data["events_total"][-1]
+            old_detectable_events = meta_data["detectable_events"][-1]
+
+            # adjust the meta data
+            # following is to keep rate the same
+            new_events_total = np.round(size*old_events_total/old_detectable_events)
+            new_total_rate = self.normalization_pdf_z * size / new_events_total
+            meta_data["events_total"][-1] = new_events_total
+            meta_data["detectable_events"][-1] = size
+            meta_data["total_rate"][-1] = new_total_rate
+
+
+            print("collected number of detectable events = ", size)
+            print("total number of events = ", new_events_total)
+            print(f"total unlensed rate (yr^-1): {new_total_rate}")
+
+            # save the meta data
+            append_json(meta_data_path, meta_data, replace=True)
             # save the final param dictionary
             append_json(output_path, param_final, replace=True)
         else:
@@ -1805,20 +1829,43 @@ class LeR(LensGalaxyParameterDistribution):
             else:
                 append_json(meta_data_path, meta_data, replace=True)
 
-            print("collected number of events = ", n)
+            print("collected number of detectable events = ", n)
             print("total number of events = ", events_total)
             print(f"total lensed rate (yr^-1): {total_rate}")
 
         print(f"storing detectable lensed params in {output_path}")
+        print(f"storing meta data in {meta_data_path}")
 
         if trim_to_size:
             # trim the final param dictionary
             print(f"\n trmming final result to size={size}")
             param_final = get_param_from_json(output_path)
+            # randomly select size number of samples
+            len_ = len(list(param_final.values())[0])
+            idx = np.random.choice(len_, size, replace=False)
[…]

### 2024-07-01 paper/paper.tex
mac

@@
 \(ler\)
-: LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers)
-Rate calculator and
+: LVK (LIGO-Virgo-KAGRA collaboration) event (compact-binary mergers)
+rate calculator and
 simulator. \textit{Journal of Open Source Software}, (), . \url{https://doi.org/10.xxxxxx/draft}}}}
@@
 \hypersetup{unicode=true,
-            pdftitle={ler : LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers) Rate calculator and simulator},
+            pdftitle={ler : LVK (LIGO-Virgo-KAGRA collaboration) event (compact-binary mergers) rate calculator and simulator},
             pdfborder={0 0 0},
@@
 
-\title{\(ler\) : LVK (LIGO-Virgo-KAGRA collaboration) Event
-(compact-binary mergers) Rate calculator and simulator}
+\title{\(ler\) : LVK (LIGO-Virgo-KAGRA collaboration) event
+(compact-binary mergers) rate calculator and simulator}
 
@@
           \author[1]{Ng Chung Yin (Leo)}
-          \author[4,5]{Justin Janquart}
+          \author[4,5,6,7]{Justin Janquart}
           \author[4,5]{Chris Van Den Broeck}
-          \author[1]{Otto Akseli HANNUKSELA}
-          \author[6]{Neha Singh}
-          \author[6]{David Keitel}
+          \author[1]{Otto Akseli Hannukkala}
+          \author[8]{Neha Singh}
+          \author[8]{David Keitel}
 
@@
 Science Park, 1098 XG Amsterdam, The Netherlands}
-      \affil[6]{Departament de F\textquotesingle isica, Universitat de
+      \affil[6]{Center for Cosmology, Particle Physics and Phenomenology
+- CP3, Universit\textquotesingle e Catholique de Louvain,
+Louvain-La-Neuve, B-1348, Belgium}
+      \affil[7]{Royal Observatory of Belgium, Avenue Circulaire, 3, 1180
+Uccle, Belgium}
+      \affil[8]{Departament de F\textquotesingle isica, Universitat de
 les Illes Balears, IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma,
@@
     SNR calculations are optimized using
-    \href{https://gwsnr.readthedocs.io/en/latest/}{\emph{gwsnr}},
-    leveraging interpolation and multiprocessing for accuracy and speed.
+    \href{https://gwsnr.readthedocs.io/en/latest/}{\emph{gwsnr}} python
+    package, leveraging interpolation and multiprocessing for accuracy
+    and speed.
   \item

### 2024-07-02 paper/paper.tex
mac

@@
 Uccle, Belgium}
-      \affil[8]{Departament de F\textquotesingle isica, Universitat de
-les Illes Balears, IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma,
-Spain}
+      \affil[8]{Departament de Física, Universitat de les Illes Balears,
+IAC3-IEEC, Crta. Valldemossa km 7.5, E-07122 Palma, Spain}
   \date{\vspace{-7ex}}
@@
 Observable GWs are created especially during the violent cosmic events
-of merging compact binaries, such as `binary black holes' (BBH), `binary
-neutron stars' (BNS), and `neutron star and black hole pair' (NSBH). GWs
-emitted by these events can be distorted or magnified by the
-gravitational fields of massive objects such as galaxies or galaxy
+of merging compact binaries, such as `binary black holes' (BBHs),
+`binary neutron stars' (BNSs), and `neutron star and black hole pair'
+(NSBHs). GWs emitted by these events can be distorted or magnified by
+the gravitational fields of massive objects such as galaxies or galaxy
 clusters, a phenomenon known as gravitational lensing. Profound
@@
 \item
-  Sampling GW Source Properties:
+  Sampling GW source properties:
 
@@
   \item
-    For the unlensed events, The sampling distribution
+    For the unlensed events, the sampling distribution
     \((\,R_m(z_s)\,)\) for the source's redshift \((\,z_s\,)\) is
@@
 \item
-  Sampling of lens galaxies attributes and source red-shifts:
+  Sampling of lens galaxy attributes and source red-shifts:
 
@@
   \item
-    For lensed case, the source redshift \((z_s)\) is sampled under the
-    strong lensing condition \((\text{SL})\), based on the precomputed
-    probability of strong lensing with source at \(z_s\)
+    For the lensed case, the source redshift \((z_s)\) is sampled under
+    the strong lensing condition \((\text{SL})\), based on the
+    precomputed probability of strong lensing with a source at \(z_s\)
     \((\text{optical depth: }P\left(\text{SL}|z_s\right) \text{ or }\tau(z_s)\,)\).
@@
     \((\theta_L)\) sampling, following Wierda et al. (2021). Rejection
-    sampling is applied on the above samples on condition that whether
-    the event is strongly lensed or not,
+    sampling is applied on the above samples depending on whether the
+    event is strongly lensed or not,
     \(P\left(\text{SL}|z_s,\theta_L\right)\).
@@
   \item
-    Source position \((\beta)\) is sampled from the caustic in the
+    The source position \((\beta)\) is sampled from the caustic in the
     source plane.
   \item
-    Sampled lens properties and source position is fed in
+    Sampled lens properties and source positions are fed to
     \emph{lenstronomy} (Birrer et al. 2021) to generate properties of
@@
     Image properties like magnification \((\mu_i)\) and time delay
-    (\(\Delta t_i\)) modifies the original source signal strength,
-    changing the signal-to-noise ratio SNR and our ability to detect.
+    (\(\Delta t_i\)) modify the original source signal strength,
+    changing the signal-to-noise ratio (SNR) and our ability to detect.
   \end{itemize}
 \item
-  Calculation of Detectable Merger Rates Per Year:
+  Calculation of detectable merger rates per year:
 
@@
 Laboratory, supported by National Science Foundation Grants
-No.~PHY-0757058 and No.~PHY-0823459.
+No.~PHY-0757058 and No.~PHY-0823459. N. Singh and D. Keitel are
+supported by Universitat de les Illes Balears (UIB); the Spanish Agencia
+Estatal de Investigación grants CNS2022-135440, PID2022-138626NB-I00,
+RED2022-134204-E, RED2022-134411-T, funded by
+MICIU/AEI/10.13039/501100011033, the European Union
+NextGenerationEU/PRTR, and the ERDF/EU; and the Comunitat Autònoma de
+les Illes Balears through the Direcció General de Recerca, Innovació I
+Transformació Digital with funds from the Tourist Stay Tax Law
+(PDR2020/11 - ITS2017-006) as well as through the Conselleria
+d'Economia, Hisenda i Innovació with grant numbers SINCO2022/6719
+(European Union NextGenerationEU/PRTR-C17.I1) and SINCO2022/18146
+(co-financed by the European Union and FEDER Operational Program
+2021-2027 of the Balearic Islands).
 

### 2024-07-03 arxiv/paper.tex
mac

--- 2024-07-02 paper/paper.tex
@@
 
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Phurailatpam Hemantakumar, Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli HANNUKSELA, Neha Singh, and David Keitel, (2024).
+\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam , Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannukkala, Neha Singh, and David Keitel, (2024).
 \(ler\)
@@
 
-        \author[1]{Phurailatpam Hemantakumar}
+        \author[1]{Hemantakumar Phurailatpam}
           \author[2,3]{Anupreeta More}
@@
 
-  {\bfseries Submitted:} 19 June 2024\\
-  {\bfseries Published:} 19 June 2024
+  {\bfseries Submitted:} 3rd July 2024\\
+  {\bfseries Published:} 3rd July 2024
 
@@
 important role in validating forthcoming lensing events (Janquart et al.
-2023).
+2023) (Collaboration et al. 2023) (Abbott et al. 2021).
 
@@
 Special thanks are extended to the academic advisors for their
-invaluable guidance and steadfast support. The collaborative efforts and
-enriching discussions with research colleagues significantly enhanced
-the study's quality. Acknowledgement is given to the Department of
-Physics, The Chinese University of Hong Kong, for the Postgraduate
-Studentship that facilitated this research. Further gratitude is
-extended to the Netherlands Organisation for Scientific Research (NWO)
-for their support. The authors also recognize the contributions of
-individuals who added empirical depth to this work. Appreciation is
-conveyed for the computational resources provided by the LIGO
-Laboratory, supported by National Science Foundation Grants
-No.~PHY-0757058 and No.~PHY-0823459. N. Singh and D. Keitel are
-supported by Universitat de les Illes Balears (UIB); the Spanish Agencia
-Estatal de Investigación grants CNS2022-135440, PID2022-138626NB-I00,
-RED2022-134204-E, RED2022-134411-T, funded by
-MICIU/AEI/10.13039/501100011033, the European Union
-NextGenerationEU/PRTR, and the ERDF/EU; and the Comunitat Autònoma de
-les Illes Balears through the Direcció General de Recerca, Innovació I
-Transformació Digital with funds from the Tourist Stay Tax Law
-(PDR2020/11 - ITS2017-006) as well as through the Conselleria
+invaluable guidance and steadfast support. Acknowledgement is given to
+the Department of Physics, The Chinese University of Hong Kong, for the
+Postgraduate Studentship that facilitated this research. Hemantakumar
+Phurailatpam and Otto A. Hannuksela acknowledge support by grants from
+the Research Grants Council of Hong Kong (Project No.~CUHK 14304622 and
+14307923), the start-up grant from the Chinese University of Hong Kong,
+and the Direct Grant for Research from the Research Committee of The
+Chinese University of Hong Kong. Further gratitude is extended to the
+Netherlands Organisation for Scientific Research (NWO) for their
+support. N. Singh and D. Keitel are supported by Universitat de les
+Illes Balears (UIB); the Spanish Agencia Estatal de Investigación grants
+CNS2022-135440, PID2022-138626NB-I00, RED2022-134204-E,
+RED2022-134411-T, funded by MICIU/AEI/10.13039/501100011033, the
+European Union NextGenerationEU/PRTR, and the ERDF/EU; and the Comunitat
+Autònoma de les Illes Balears through the Direcció General de Recerca,
+Innovació I Transformació Digital with funds from the Tourist Stay Tax
+Law (PDR2020/11 - ITS2017-006) as well as through the Conselleria
 d'Economia, Hisenda i Innovació with grant numbers SINCO2022/6719
@@
 (co-financed by the European Union and FEDER Operational Program
-2021-2027 of the Balearic Islands).
+2021-2027 of the Balearic Islands). The authors also recognize the
+contributions of individuals who added empirical depth to this work.
+Appreciation is conveyed for the computational resources provided by the
+LIGO Laboratory, supported by National Science Foundation Grants
+No.~PHY-0757058 and No.~PHY-0823459.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Abbott2021}
+Abbott, R., T. D. Abbott, S. Abraham, F. Acernese, K. Ackley, A. Adams,
+C. Adams, et al. 2021. {``Search for Lensing Signatures in the
+Gravitational-Wave Observations from the First Half of LIGO--Virgo's
+Third Observing Run.''} \emph{The Astrophysical Journal} 923 (1): 14.
+\url{https://doi.org/10.3847/1538-4357/ac23db}.
+
 \bibitem[\citeproctext]{ref-VIRGO2015}
@@
 \url{https://doi.org/10.21105/joss.03283}.
+
+\bibitem[\citeproctext]{ref-ligolensing2023}
+Collaboration, The LIGO Scientific, the Virgo Collaboration, the KAGRA
+Collaboration, R. Abbott, H. Abe, F. Acernese, K. Ackley, et al. 2023.
+{``Search for Gravitational-Lensing Signatures in the Full Third
+Observing Run of the LIGO-Virgo Network.''}
+\url{https://arxiv.org/abs/2304.08393}.
 

### 2024-07-03 ler/lens_galaxy_population/jit_functions.py

@@ -228,6 +228,7 @@ def lens_redshift_SDSS_catalogue(zs, splineDc, splineDcInv, u, cdf):
     splineDcInv: `list`
         List of spline coefficients for the inverse of comoving distance and redshifts
     u: `numpy.ndarray` (1D array of float of size=size)
+        corresponding x values wrt to the cdf values
         e.g. u = np.linspace(0, 1, 500)
     cdf: `numpy.ndarray` (1D array of float of size=size)
         Cumulative distribution function of the lens redshift distribution between 0 and 1

### 2024-07-03 ler/lens_galaxy_population/lens_galaxy_parameter_distribution.py

@@ -601,7 +601,7 @@ class LensGalaxyParameterDistribution(CBCSourceParameterDistribution, ImagePrope
                 size=size)
 
             # sample gravitional waves source parameter
-            param = dict(zs=np.array(zs))
+            param = dict(zs=lens_parameters["zs"])
             if samplers_params["source_parameters"]:
                 param.update(self.sample_gw_parameters(size=size))
             gw_param = self.sample_source_parameters(size=size, param=param)

### 2024-07-06 arxiv/paper.tex
mac

@@
 rate calculator and
-simulator. \textit{Journal of Open Source Software}, (), . \url{https://doi.org/10.xxxxxx/draft}}}}
+simulator. \textit{PendingJournal of Open Source Software}, Pending(Pending), Pending. \url{https://doi.org/10.xxxxxx/draft}}}}
 
@@
   \setlength\itemsep{0em}
-    \item \href{https://github.com/@Pending Reviewer}{@@Pending Reviewer}
+    \item \href{https://github.com/Pending Reviewer}{@Pending Reviewer}
     \item \href{https://github.com/}{@}
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
 
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-Abbott2021}
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Abbott2021}
 Abbott, R., T. D. Abbott, S. Abraham, F. Acernese, K. Ackley, A. Adams,
@@
 
-\bibitem[\citeproctext]{ref-VIRGO2015}
+\leavevmode\hypertarget{ref-VIRGO2015}
 Acernese, F, M Agathos, K Agatsuma, D Aisa, N Allemandou, A Allocca, J
@@
 
-\bibitem[\citeproctext]{ref-KAGRA2021}
+\leavevmode\hypertarget{ref-KAGRA2021}
 Akutsu, T, M Ando, K Arai, Y Arai, S Araki, A Araya, N Aritomi, et al.
@@
 
-\bibitem[\citeproctext]{ref-Ashton2019}
+\leavevmode\hypertarget{ref-Ashton2019}
 Ashton, Gregory, Moritz Hübner, Paul D. Lasky, Colm Talbot, Kendall
@@
 
-\bibitem[\citeproctext]{ref-astropy}
+\leavevmode\hypertarget{ref-astropy}
 Astropy Collaboration. 2013. {``{Astropy: A community Python package for
@@
 
-\bibitem[\citeproctext]{ref-Birrer2021}
+\leavevmode\hypertarget{ref-Birrer2021}
 Birrer, Simon, Anowar J. Shajib, Daniel Gilman, Aymeric Galan, Jelle
@@
 
-\bibitem[\citeproctext]{ref-ligolensing2023}
+\leavevmode\hypertarget{ref-ligolensing2023}
 Collaboration, The LIGO Scientific, the Virgo Collaboration, the KAGRA
@@
 
-\bibitem[\citeproctext]{ref-gwcosmo}
+\leavevmode\hypertarget{ref-gwcosmo}
 gwcosmo Contributors. 2022. {``{gwcosmo: A Python package for
@@
 
-\bibitem[\citeproctext]{ref-Janquart2023}
+\leavevmode\hypertarget{ref-Janquart2023}
 Janquart, J, M Wright, S Goyal, J C L Chan, A Ganguly, Á Garrón, D
@@
 
-\bibitem[\citeproctext]{ref-numba}
+\leavevmode\hypertarget{ref-numba}
 Lam, Stan, Stéphane Pitrou, and Mark Seibert. 2022. {``Numba: A High
@@
 
-\bibitem[\citeproctext]{ref-numpy}
+\leavevmode\hypertarget{ref-numpy}
 NumPy Community. 2022. {``NumPy: A Fundamental Package for Scientific
@@
 
-\bibitem[\citeproctext]{ref-LIGO2015}
+\leavevmode\hypertarget{ref-LIGO2015}
 The LIGO Scientific Collaboration, J Aasi, B P Abbott, R Abbott, T
@@
 
-\bibitem[\citeproctext]{ref-scipy}
+\leavevmode\hypertarget{ref-scipy}
 Virtanen, Pauli, Ralf Gommers, Travis E. Oliphant, Matt Haberland, Tyler
@@
 
-\bibitem[\citeproctext]{ref-Wempe2022}
+\leavevmode\hypertarget{ref-Wempe2022}
 Wempe, Ewoud, Léon V. E. Koopmans, A. Renske A. C. Wierda, Otto Akseli
@@
 
-\bibitem[\citeproctext]{ref-Wierda2021}
+\leavevmode\hypertarget{ref-Wierda2021}
 Wierda, A. Renske A. C., Ewoud Wempe, Otto A. Hannuksela, Léon V. E.
@@
 
-\end{CSLReferences}
-
 \end{document}

### 2024-07-06 paper/paper.tex
mac

@@
     Intrinsic and extrinsic parameters \((\theta)\) of GW sources are
-    sampled using prior distributions \((P(\theta))\) from the gwcosmo
-    (gwcosmo Contributors 2022) and bilby (Ashton et al. 2019) packages,
-    with options for users to input custom distributions.
+    sampled using prior distributions \((P(\theta))\) from the
+    \emph{gwcosmo} (gwcosmo Contributors 2022) and \emph{bilby} (Ashton
+    et al. 2019) packages, with options for users to input custom
+    distributions.
   \end{itemize}
@@
   \item
-    The package utilizes the Elliptical Power Law with external shear
-    (EPL+Shear) model (Wempe et al. 2022) for galaxy parameters
-    \((\theta_L)\) sampling, following Wierda et al. (2021). Rejection
+    Following Wierda et al. (2021), the package utilizes the Elliptical
+    Power Law with external shear (EPL+Shear) model (Wempe et al. 2022)
+    for sampling the galaxy parameters \((\theta_L)\). Rejection
     sampling is applied on the above samples depending on whether the

### 2024-07-29 ler/image_properties/image_properties.py
mac

@@ -67,7 +67,7 @@ class ImageProperties():
         default: "./interpolator_pickle"
     create_new_interpolator : `dict`
         dictionary to create new interpolator pickle files
-        default: dict(Dl_to_z=dict(create_new=False, resolution=500))
+        default: dict(Dl_to_z=dict(create_new=False, resolution=1000))
 
     Examples
     --------
@@ -140,13 +140,13 @@ class ImageProperties():
         
         # initialize the interpolator's parameters
         self.create_new_interpolator = dict(
-            Dl_to_z=dict(create_new=False, resolution=500),
+            Dl_to_z=dict(create_new=False, resolution=1000),
         )
         if isinstance(create_new_interpolator, dict):
             self.create_new_interpolator.update(create_new_interpolator)
         elif create_new_interpolator is True:
             self.create_new_interpolator = dict(
-                Dl_to_z=dict(create_new=True, resolution=500)
+                Dl_to_z=dict(create_new=True, resolution=1000)
             )
 
         resolution = self.create_new_interpolator["Dl_to_z"]["resolution"]

### 2024-07-29 ler/lens_galaxy_population/lens_galaxy_parameter_distribution.py

@@ -233,31 +233,31 @@ class LensGalaxyParameterDistribution(CBCSourceParameterDistribution, ImagePrope
         self.directory = directory
         # initialize the interpolator's parameters
         self.create_new_interpolator = dict(
-            redshift_distribution=dict(create_new=False, resolution=500),
-            z_to_luminosity_distance=dict(create_new=False, resolution=500),
-            velocity_dispersion=dict(create_new=False, resolution=500),
-            axis_ratio=dict(create_new=False, resolution=500),
-            optical_depth=dict(create_new=False, resolution=100),
-            z_to_Dc=dict(create_new=False, resolution=500),
-            Dc_to_z=dict(create_new=False, resolution=500),
-            angular_diameter_distance=dict(create_new=False, resolution=500),
-            differential_comoving_volume=dict(create_new=False, resolution=500),
-            Dl_to_z=dict(create_new=False, resolution=500),
+            redshift_distribution=dict(create_new=False, resolution=1000),
+            z_to_luminosity_distance=dict(create_new=False, resolution=1000),
+            velocity_dispersion=dict(create_new=False, resolution=1000),
+            axis_ratio=dict(create_new=False, resolution=1000),
+            optical_depth=dict(create_new=False, resolution=200),
+            z_to_Dc=dict(create_new=False, resolution=1000),
+            Dc_to_z=dict(create_new=False, resolution=1000),
+            angular_diameter_distance=dict(create_new=False, resolution=1000),
+            differential_comoving_volume=dict(create_new=False, resolution=1000),
+            Dl_to_z=dict(create_new=False, resolution=1000),
         )
         if isinstance(create_new_interpolator, dict):
             self.create_new_interpolator.update(create_new_interpolator)
         elif create_new_interpolator is True:
             self.create_new_interpolator = dict(
-                redshift_distribution=dict(create_new=True, resolution=500),
-                z_to_luminosity_distance=dict(create_new=True, resolution=500),
-                velocity_dispersion=dict(create_new=True, resolution=500),
-                axis_ratio=dict(create_new=True, resolution=500),
-                optical_depth=dict(create_new=True, resolution=100),
-                z_to_Dc=dict(create_new=True, resolution=500),
-                Dc_to_z=dict(create_new=True, resolution=500),
-                angular_diameter_distance=dict(create_new=True, resolution=500),
-                differential_comoving_volume=dict(create_new=True, resolution=500),
-                Dl_to_z=dict(create_new=True, resolution=500),
[…]

### 2024-07-29 ler/lens_galaxy_population/optical_depth.py

@@ -186,25 +186,25 @@ class OpticalDepth():
 
         self.directory = directory
         self.c_n_i = dict(
-            velocity_dispersion=dict(create_new=False, resolution=500),
-            axis_ratio=dict(create_new=False, resolution=500),
-            optical_depth=dict(create_new=False, resolution=100),
-            z_to_Dc=dict(create_new=False, resolution=500),
-            Dc_to_z=dict(create_new=False, resolution=500),
-            angular_diameter_distance=dict(create_new=False, resolution=500),
-            differential_comoving_volume=dict(create_new=False, resolution=500),
+            velocity_dispersion=dict(create_new=False, resolution=1000),
+            axis_ratio=dict(create_new=False, resolution=1000),
+            optical_depth=dict(create_new=False, resolution=200),
+            z_to_Dc=dict(create_new=False, resolution=1000),
+            Dc_to_z=dict(create_new=False, resolution=1000),
+            angular_diameter_distance=dict(create_new=False, resolution=1000),
+            differential_comoving_volume=dict(create_new=False, resolution=1000),
         )
         if isinstance(create_new_interpolator, dict):
             self.c_n_i.update(create_new_interpolator)
         elif create_new_interpolator is True:
             self.c_n_i = dict(
-                velocity_dispersion=dict(create_new=True, resolution=500),
-                axis_ratio=dict(create_new=True, resolution=500),
-                optical_depth=dict(create_new=True, resolution=100),
-                z_to_Dc=dict(create_new=True, resolution=500),
-                Dc_to_z=dict(create_new=True, resolution=500),
-                angular_diameter_distance=dict(create_new=True, resolution=500),
-                differential_comoving_volume=dict(create_new=True, resolution=500),
+                velocity_dispersion=dict(create_new=True, resolution=1000),
+                axis_ratio=dict(create_new=True, resolution=1000),
+                optical_depth=dict(create_new=True, resolution=200),
+                z_to_Dc=dict(create_new=True, resolution=1000),
+                Dc_to_z=dict(create_new=True, resolution=1000),
+                angular_diameter_distance=dict(create_new=True, resolution=1000),
+                differential_comoving_volume=dict(create_new=True, resolution=1000),
             )
 
         vd_name = self.sampler_priors["velocity_dispersion"]  # velocity dispersion sampler name

### 2024-07-29 ler/rates/gwrates.py

@@ -3,10 +3,10 @@
 This module contains the main class for calculating the rates of detectable gravitational waves events. The class inherits the :class:`~ler.gw_source_population.CBCSourceParameterDistribution` class for source parameters sampling and uses `gwsnr` package for SNR calculation. 
 """
 
-import contextlib
 import os
 import warnings
 warnings.filterwarnings("ignore")
+import contextlib
 import numpy as np
 from scipy.stats import norm
 from astropy.cosmology import LambdaCDM

### 2024-07-29 ler/rates/ler.py

@@ -15,7 +15,7 @@ from ..utils import load_json, append_json, get_param_from_json, batch_handler, 
 
 
 class LeR(LensGalaxyParameterDistribution):
-    """Class to calculate both the rates of lensed and unlensed events. Please note that parameters of the simulated events are stored in json file but not as an attribute of the class. This saves RAM memory. 
+    """Class to sample of lensed and unlensed events and calculate it's rates. Please note that parameters of the simulated events are stored in json file but not as an attribute of the class. This saves RAM memory. 
 
     Parameters
     ----------
@@ -29,25 +29,54 @@ class LeR(LensGalaxyParameterDistribution):
     z_max : `float`
         maximum redshift.
         default z_max = 10.
-        for popI_II, popIII, primordial, BNS z_max = 10., 40., 40., 2. respectively.
+        for popI_II, popIII, primordial, BNS z_max = 10., 40., 40., 5. respectively.
+    event_type : `str`
+        type of event to generate.
+        default event_type = 'BBH'. Other options are 'BNS', 'NSBH'.
     size : `int`
         number of samples for sampling.
-        default size = 100000.
+        default size = 100000. To get stable rates, size should be large (>=1e6).
     batch_size : `int`
         batch size for SNR calculation.
         default batch_size = 50000.
         reduce the batch size if you are getting memory error.
-        recommended batch_size = 50000, if size = 1000000.
-    snr_finder : `str`
+        recommended batch_size = 200000, if size = 1000000.
+    cosmology : `astropy.cosmology`
+        cosmology to use for the calculation.
+        default cosmology = LambdaCDM(H0=70, Om0=0.3, Ode0=0.7).
+    snr_finder : `str` or `function`
         default snr_finder = 'gwsnr'.
         if None, the SNR will be calculated using the gwsnr package.
-        if 'custom', the SNR will be calculated using a custom function.
-        The custom function should have input and output as given in GWSNR.snr method.
+        if custom snr finder function is provided, the SNR will be calculated using a custom function. The custom function should follow the following signature:
+        def snr_finder(gw_param_dict):
+            ...
+            return optimal_snr_dict
+        where optimal_snr_dict.keys = ['optimal_snr_net']. Refer to `gwsnr` package's GWSNR.snr attribute for more details.
+    pdet_finder : `function`
+        default pdet_finder = None.
[…]

### 2024-07-29 paper/arxiv/paper.tex
mac

--- 2024-07-06 arxiv/paper.tex
@@
 
-\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam , Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannukkala, Neha Singh, and David Keitel, (2024).
+\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam , Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannuksela, Neha Singh, and David Keitel, (2024).
 \(ler\)
@@
           \author[4,5]{Chris Van Den Broeck}
-          \author[1]{Otto Akseli Hannukkala}
+          \author[1]{Otto Akseli Hannuksela}
           \author[8]{Neha Singh}

### 2024-07-29 ler/__init__.py

@@ -14,7 +14,7 @@ except RuntimeError:
 __author__ = 'hemanta_ph <hemantaphurailatpam@gmail.com>'
 
 # The version as used in the setup.py
-__version__ = "0.3.7"
+__version__ = "0.3.8"
 
 # add __file__
 import os

### 2024-07-29 ler/gw_source_population/cbc_source_parameter_distribution.py

@@ -291,27 +291,21 @@ class CBCSourceParameterDistribution(CBCSourceRedshiftDistribution):
             event_type, source_priors, source_priors_params
         )
 
-        if self.gw_param_samplers["zs"] == "sample_source_redshift":
-            # initialize the SourceGalaxyPopulationModel mother class
-            # for redshift distribution
-            # instance attribute sample_source_redshift is initialized here
-            super().__init__(
-                z_min=z_min,
-                z_max=z_max,
-                event_type=event_type,
-                merger_rate_density=self.gw_param_samplers["merger_rate_density"],
-                merger_rate_density_param=self.gw_param_samplers_params[
-                    "merger_rate_density"
-                ],
-                cosmology=cosmology,
-                directory=directory,
-                create_new_interpolator=self.create_new_interpolator,
-            )
-        else:
-            # if you already have the redshift distribution function, you can pass it in.
-            # super class will not be initialized anymore,
-            # get z_to_luminosity_distance
-            self.lookup_table_luminosity_distance(z_min, z_max, directory)
+        # initialize the SourceGalaxyPopulationModel mother class
+        # for redshift distribution
+        # instance attribute sample_source_redshift is initialized here
+        super().__init__(
+            z_min=z_min,
+            z_max=z_max,
+            event_type=event_type,
+            merger_rate_density=self.gw_param_samplers["merger_rate_density"],
+            merger_rate_density_param=self.gw_param_samplers_params[
+                "merger_rate_density"
+            ],
+            cosmology=cosmology,
+            directory=directory,
+            create_new_interpolator=self.create_new_interpolator,
+        )
 
         # initializing samplers
         # it goes through the setter functions and assign the sampler functions

### 2024-07-29 ler/gw_source_population/cbc_source_redshift_distribution.py

@@ -53,7 +53,7 @@ class CBCSourceRedshiftDistribution(object):
         default: './interpolator_pickle'
     create_new_interpolator : `dict`
         Dictionary of interpolator creation parameters
-        default: None/dict(redshift_distribution=dict(create_new=False, resolution=500), z_to_luminosity_distance=dict(create_new=False, resolution=500), differential_comoving_volume=dict(create_new=False, resolution=500))
+        default: None/dict(redshift_distribution=dict(create_new=False, resolution=1000), z_to_luminosity_distance=dict(create_new=False, resolution=1000), differential_comoving_volume=dict(create_new=False, resolution=1000))
 
     Examples
     ----------
@@ -170,7 +170,7 @@ class CBCSourceRedshiftDistribution(object):
     c_n_i = None
     """``dict`` \n
     c_n_i stands for 'create new interpolator'. Dictionary of interpolator creation parameters. \n
-    e.g. dict(redshift_distribution=dict(create_new=False, resolution=500), z_to_luminosity_distance=dict(create_new=False, resolution=500), differential_comoving_volume=dict(create_new=False, resolution=500))
+    e.g. dict(redshift_distribution=dict(create_new=False, resolution=1000), z_to_luminosity_distance=dict(create_new=False, resolution=1000), differential_comoving_volume=dict(create_new=False, resolution=1000))
     """
 
     normalization_pdf_z = None
@@ -207,7 +207,7 @@ class CBCSourceRedshiftDistribution(object):
         self.cosmo = cosmology if cosmology else LambdaCDM(H0=70, Om0=0.3, Ode0=0.7)
         # setting up the interpolator creation parameters
         self.c_n_i = dict(
-            redshift_distribution=dict(create_new=False, resolution=500), z_to_luminosity_distance=dict(create_new=False, resolution=500), differential_comoving_volume=dict(create_new=False, resolution=500))
+            redshift_distribution=dict(create_new=False, resolution=1000), z_to_luminosity_distance=dict(create_new=False, resolution=1000), differential_comoving_volume=dict(create_new=False, resolution=1000))
         if create_new_interpolator:
             self.c_n_i.update(create_new_interpolator)
         # creating of interpolators for redshift dependent quantities

### 2024-07-29 ler/utils/utils.py

@@ -69,6 +69,19 @@ def load_json(file_name):
 
     return param
 
+def save_json(file_name, param):
+    """Save a dictionary as a json file.
+
+    Parameters
+    ----------
+    file_name : `str`
+        json file name for storing the parameters.
+    param : `dict`
+        dictionary to be saved as a json file.
+    """
+    with open(file_name, "w", encoding="utf-8") as write_file:
+        json.dump(param, write_file)
+
 def append_json(file_name, new_dictionary, old_dictionary=None, replace=False):
     """Append and update a json file with a dictionary.
 
@@ -237,7 +250,6 @@ def rejection_sample2d(pdf, xmin, xmax, ymin, ymax, size=100, chunk_size=10000):
     x_sample : `numpy.ndarray`
         samples from the pdf in the x-axis.
     """
-    chunk_size = 10000
 
     x = np.random.uniform(xmin, xmax, chunk_size)
     y = np.random.uniform(ymin, ymax, chunk_size)

### 2024-07-29 setup.py

@@ -13,7 +13,7 @@ if python_version < (3, 10):
 
 setup(
     name="ler",
-    version="0.3.7",
+    version="0.3.8",
     description="Gravitational waves Lensing Rates",
     long_description=long_description,
     long_description_content_type='text/markdown',
@@ -31,7 +31,7 @@ setup(
         "numba>=0.57.1",
         "bilby>=1.0.2",
         "gwsnr>=0.2.0",
-        "scipy>=1.9.0",
+        "scipy<1.14.0",
         "lenstronomy>=1.10.4",
         "astropy>=5.1",
         "tqdm>=4.64.1",

### 2024-08-20 ler/rates/ler.py
mac

@@ -11,7 +11,7 @@ import numpy as np
 from scipy.stats import norm
 from astropy.cosmology import LambdaCDM
 from ..lens_galaxy_population import LensGalaxyParameterDistribution
-from ..utils import load_json, append_json, get_param_from_json, batch_handler, add_dict_values
+from ..utils import load_json, append_json, get_param_from_json, batch_handler
 
 
 class LeR(LensGalaxyParameterDistribution):

### 2024-08-20 ler/utils/utils.py
mac

@@ -143,25 +143,25 @@ def append_json(file_name, new_dictionary, old_dictionary=None, replace=False):
 
     return data
 
-def add_dict_values(dict1, dict2):
-    """Adds the values of two dictionaries together.
+# def add_dict_values(dict1, dict2):
+#     """Adds the values of two dictionaries together.
     
-    Parameters
-    ----------
-    dict1 : `dict`
-        dictionary to be added.
-    dict2 : `dict`
-        dictionary to be added.
+#     Parameters
+#     ----------
+#     dict1 : `dict`
+#         dictionary to be added.
+#     dict2 : `dict`
+#         dictionary to be added.
 
-    Returns
-    ----------
-    dict1 : `dict`
-        dictionary with added values.
-    """
-    data_key = dict1.keys()
-    for key, value in dict2.items():
-        if key in data_key:
-            dict1[key] = np.concatenate((dict1[key], value))
+#     Returns
+#     ----------
+#     dict1 : `dict`
+#         dictionary with added values.
+#     """
+#     data_key = dict1.keys()
+#     for key, value in dict2.items():
+#         if key in data_key:
+#             dict1[key] = np.concatenate((dict1[key], value))
 
     return dict1
 
@@ -370,9 +370,9 @@ def create_func_pdf_invcdf(x, y, category="function"):
     Parameters
     ----------
     x : `numpy.ndarray`
-        x values.
+        x values. This has to sorted in ascending order.
     y : `numpy.ndarray`
-        y values.
+        y values. Corresponding to the x values.
     category : `str`, optional
         category of the function. Default is "function". Other options are "function_inverse", "pdf" and "inv_cdf".
 
